Reproducible Research 
=====================
### Peer Assessment 1
### *Alexis Segovia*
### 17/12/2015

Used libraries:

```{r, echo = TRUE}
library(ggplot2)
library(scales)
library(Hmisc)
library(sqldf)
library(plyr)
```
### Loading and preprocessing the data

1. Load the data (i.e. read.csv())

```{r,echo=TRUE}
activityData <- read.csv('activity.csv')
```

2. Process/transform the data (if necessary) into a format suitable for your analysis

```{r,echo=TRUE}
activityData$date <- as.Date(activityData$date)
```

### What is mean total number of steps taken per day?

```{r,echo=TRUE}
stepsTakenPerDay <- tapply(activityData$steps, activityData$date, sum, na.rm=TRUE)
```

1. Make a histogram of the total number of steps taken each day

```{r,echo=TRUE}
qplot(stepsTakenPerDay,geom="histogram",binwidth = 400,main ='Total steps taken per day', xlab="Steps", ylab='Frequency', fill=I("blue"),  col=I("red")) 
```

2. Calculate and report the mean and median total number of steps taken per day

```{r,echo=TRUE}
stepsTakenPerDayMean <- mean(stepsTakenPerDay)
stepsTakenPerDayMedian <- median(stepsTakenPerDay)
```

- Mean  : `r stepsTakenPerDayMean`
- Median: `r stepsTakenPerDayMedian`

### What is the average daily activity pattern?

```{r,echo=TRUE}
averDailyActivity <- aggregate(x=list(meanSteps=activityData$steps), by=list(interval=activityData$interval), FUN=mean, na.rm=TRUE)
```

1. Make a time series plot

```{r,echo=TRUE}
ggplot(data=averDailyActivity, aes(x=interval, y=meanSteps)) +
        geom_line() + xlab("5-minute interval") + ylab("average number of steps taken") 
```

2. Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?

```{r,echo=TRUE}
StepsByInterval <- aggregate(steps ~ interval, activityData, mean)
plot(StepsByInterval$interval,StepsByInterval$steps, type="l", xlab="Interval", ylab="Number of Steps",main="Average Number of Steps per Day by Interval")
maxNUmberSteps <- StepsByInterval[which.max(StepsByInterval$steps),1]
```
Max Number of Steps : `r maxNUmberSteps`

### Imputing missing values

1. Calculate and report the total number of missing values in the dataset (i.e. the total number of rows with NAs)

```{r,echo=TRUE}
totalNA <- sqldf('SELECT ad.* FROM "activityData" as ad
    WHERE ad.steps IS NULL ORDER BY ad.date, ad.interval ') 
```

Total Number of missing values: `r NROW(totalNA)` 

2. Devise a strategy for filling in all of the missing values in the dataset. The strategy does not need to be sophisticated. For example, you could use the mean/median for that day, or the mean for that 5-minute interval, etc.

```{r,echo=TRUE}
activityDataNA <- na.omit(activityData)
intervals <- ddply(activityDataNA,~interval, summarise, mean=mean(steps))
results <- sqldf('SELECT ad.*, ada.mean  FROM "intervals" as ada
    JOIN "activityData" as ad
    ON ad.interval = ada.interval 
    ORDER BY ad.date, ad.interval ') 

```

3. Create a new dataset that is equal to the original dataset but with the missing data filled in.

```{r,echo=TRUE}
results$steps[is.na(results$steps)] <- results$mean[is.na(results$steps)]
```

4. Make a histogram of the total number of steps taken each day and Calculate and report the mean and median total number of steps taken per day.

Prepare data to plot

```{r,echo=TRUE}
ResultsTotalSteps <- as.integer( sqldf(' 
    SELECT sum(steps)  
    FROM results') )

ResultsTotalStepsDaily <- sqldf(' 
    SELECT date, sum(steps) as "ResultsTotalStepsDaily" 
    FROM results GROUP BY date 
    ORDER BY date') 

Resultsdailysteps <- sqldf('   
    SELECT date, ResultsTotalStepsDaily as "steps"
    FROM "ResultsTotalStepsDaily"
    ORDER BY date') 
```

Histogram of the total number of steps taken each day.

```{r,echo=TRUE}
hist(Resultsdailysteps$steps, 
     main="After Imputate NA",
     breaks=10,
     xlab="Total Number of Steps Taken Daily")
```

### Are there differences in activity patterns between weekdays and weekends?

Adding a column called (Week).

```{r,echo=TRUE}
results$week <- as.factor(ifelse(weekdays(results$date) %in% 
                                         c("sabado","domingo"),"weekend", "weekday"))


ResultsWeek <- sqldf('   
    SELECT interval, avg(steps) as "AverageSteps", week
    FROM results
    GROUP BY week, interval
    ORDER BY interval ')
```


```{r,echo=TRUE}
p <- xyplot(AverageSteps ~ interval | factor(week), data=ResultsWeek, 
            type = 'l',
            main="Differences in activity patterns between weekdays and weekends",
            xlab="5-Minute Interval",
            ylab="Average of Steps Taken")
print (p) 
```

*Yes. There are more activities at weekend*